% bibliography_supp.tex --- SUPPLEMENT reference list (full).
% Entries [1]--[28] repeat bibliography.tex verbatim (same order) so a citation
% number appearing in both documents refers to the same source; entries [29]--[45]
% are supplement-only, in supplement first-citation order (S1--S4, then Appendices A--F).

\begin{thebibliography}{99}

\bibitem{Pittman1995PRA}
T.~B.~Pittman, Y.~H.~Shih, D.~V.~Strekalov, and A.~V.~Sergienko, ``Optical imaging by means of two-photon quantum entanglement,'' \emph{Phys. Rev. A} \textbf{52}, R3429--R3432 (1995).

\bibitem{Bennink2002PRL}
R.~S.~Bennink, S.~J.~Bentley, and R.~W.~Boyd, ``\,`Two-Photon' Coincidence Imaging with a Classical Source,'' \emph{Phys. Rev. Lett.} \textbf{89}, 113601 (2002).

\bibitem{Shapiro2008PRA}
J.~H.~Shapiro, ``Computational ghost imaging,'' \emph{Phys. Rev. A} \textbf{78}, 061802(R) (2008).

\bibitem{Bromberg2009PRA}
Y.~Bromberg, O.~Katz, and Y.~Silberberg, ``Ghost imaging with a single detector,'' \emph{Phys. Rev. A} \textbf{79}, 053840 (2009).

\bibitem{Bina2013PRL}
M.~Bina \emph{et al.}, ``Backscattering Differential Ghost Imaging in Turbid Media,'' \emph{Phys. Rev. Lett.} \textbf{110}, 083901 (2013).

\bibitem{Cheng2009OE}
J.~Cheng, ``Ghost imaging through turbulent atmosphere,'' \emph{Opt. Express} \textbf{17}(10), 7916--7921 (2009).

\bibitem{PengChen2024APL}
Y.~Peng and W.~Chen, ``Deep learning-enhanced ghost imaging through dynamic and complex scattering media with supervised corrections of dynamic scaling factors,'' \emph{Appl. Phys. Lett.} \textbf{124}, 181104 (2024).

\bibitem{PengXiaoChen2025OE}
Y.~Peng, Y.~Xiao, and W.~Chen, ``High-fidelity optical wireless transmission in complex environments around a corner using the design of a single-layer neural network for data encoding,'' \emph{Opt. Express} \textbf{33}(14), 30123--30135 (2025).

\bibitem{Duarte2008SPM}
M.~F.~Duarte \emph{et al.}, ``Single-Pixel Imaging via Compressive Sampling,'' \emph{IEEE Signal Process. Mag.} \textbf{25}(2), 83--91 (2008).

\bibitem{Edgar2019NatPhoton}
M.~P.~Edgar, G.~M.~Gibson, and M.~J.~Padgett, ``Principles and prospects for single-pixel imaging,'' \emph{Nat. Photonics} \textbf{13}, 13--20 (2019).

\bibitem{Zhang2017OE}
Z.~Zhang, X.~Wang, G.~Zheng, and J.~Zhong, ``Hadamard single-pixel imaging versus Fourier single-pixel imaging,'' \emph{Opt. Express} \textbf{25}(16), 19619--19639 (2017).

\bibitem{Yu2019Sensors}
W.-K.~Yu, ``Super Sub-Nyquist Single-Pixel Imaging by Means of Cake-Cutting Hadamard Basis Sort,'' \emph{Sensors} \textbf{19}(19), 4122 (2019).

\bibitem{KechKrahmer1603.07316}
M.~Kech and F.~Krahmer, ``Optimal Injectivity Conditions for Bilinear Inverse Problems with Applications to Identifiability of Deconvolution Problems,'' \emph{SIAM J. Appl. Algebra Geom.} \textbf{1}(1), 20--37 (2017).

\bibitem{LiLeeBresler1501.06120}
Y.~Li, K.~Lee, and Y.~Bresler, ``Identifiability in Bilinear Inverse Problems With Applications to Subspace or Sparsity-Constrained Blind Gain and Phase Calibration,'' \emph{IEEE Trans. Inf. Theory} \textbf{63}(2), 822--842 (2017).

\bibitem{AhmedRechtRomberg1211.5608}
A.~Ahmed, B.~Recht, and J.~Romberg, ``Blind Deconvolution Using Convex Programming,'' \emph{IEEE Trans. Inf. Theory} \textbf{60}(3), 1711--1732 (2014).

\bibitem{ChoudharyMitra1402.2637}
S.~Choudhary and U.~Mitra, ``Identifiability Scaling Laws in Bilinear Inverse Problems,'' arXiv:1402.2637 (2014).

\bibitem{KuengRauhutTerstiege1410.6913}
R.~Kueng, H.~Rauhut, and U.~Terstiege, ``Low rank matrix recovery from rank one measurements,'' \emph{Appl. Comput. Harmon. Anal.} \textbf{42}(1), 88--116 (2017).

\bibitem{Ferri2010PRL}
F.~Ferri, D.~Magatti, L.~A.~Lugiato, and A.~Gatti, ``Differential Ghost Imaging,'' \emph{Phys. Rev. Lett.} \textbf{104}, 253603 (2010).

\bibitem{Sun2012OE}
B.~Sun \emph{et al.}, ``Normalized ghost imaging,'' \emph{Opt. Express} \textbf{20}(15), 16892--16901 (2012).

\bibitem{WeissFriedlander1990}
A.~J.~Weiss and B.~Friedlander, ``Eigenstructure methods for direction finding with sensor gain and phase uncertainties,'' \emph{Circuits Syst. Signal Process.} \textbf{9}(3), 271--300 (1990).

\bibitem{Balzano2007IPSN}
L.~Balzano and R.~Nowak, ``Blind calibration of sensor networks,'' in \emph{Proc. IPSN}, 2007, pp.~79--88.

\bibitem{LingStrohmer2015}
S.~Ling and T.~Strohmer, ``Self-calibration and biconvex compressive sensing,'' \emph{Inverse Problems} \textbf{31}(11), 115002 (2015).

\bibitem{Nie2021PRA}
X.~Nie \emph{et al.}, ``Noise-robust computational ghost imaging with pink noise speckle patterns,'' \emph{Phys. Rev. A} \textbf{104}, 013513 (2021).

\bibitem{Han2025OE}
Y.~Han \emph{et al.}, ``Generalized sparse Hadamard single-pixel imaging,'' \emph{Opt. Express} \textbf{33}(20), 43324--43341 (2025).

\bibitem{Tsybakov2009}
A.~B.~Tsybakov, \emph{Introduction to Nonparametric Estimation}, Springer Series in Statistics. New York: Springer, 2009.

\bibitem{GillLevit1995}
R.~D.~Gill and B.~Y.~Levit, ``Applications of the van Trees inequality: a Bayesian Cram\'er--Rao bound,'' \emph{Bernoulli} \textbf{1}(1--2), 59--79 (1995).

\bibitem{Pinsker1980}
M.~S.~Pinsker, ``Optimal filtering of square-integrable signals in Gaussian noise,'' \emph{Probl. Inf. Transm.} \textbf{16}(2), 120--133 (1980).

\bibitem{Anscombe1948}
F.~J.~Anscombe, ``The transformation of Poisson, binomial and negative-binomial data,'' \emph{Biometrika} \textbf{35}(3--4), 246--254 (1948).

\bibitem{BrownForsythe1974}
M.~B.~Brown and A.~B.~Forsythe, ``Robust tests for the equality of variances,'' \emph{J. Amer. Statist. Assoc.} \textbf{69}(346), 364--367 (1974).

\bibitem{Levene1960}
H.~Levene, ``Robust tests for equality of variances,'' in \emph{Contributions to Probability and Statistics: Essays in Honor of Harold Hotelling}, I.~Olkin \emph{et al.}, Eds. Stanford, CA: Stanford Univ. Press, 1960, pp.~278--292.

\bibitem{Serfling1974}
R.~J.~Serfling, ``Probability inequalities for the sum in sampling without replacement,'' \emph{Ann. Statist.} \textbf{2}(1), 39--48 (1974).

\bibitem{Davydov1968}
Yu.~A.~Davydov, ``Convergence of distributions generated by stationary stochastic processes,'' \emph{Theory Probab. Appl.} \textbf{13}(4), 691--696 (1968).

\bibitem{MerlevedePeligradRio2011}
F.~Merlev\`ede, M.~Peligrad, and E.~Rio, ``A Bernstein type inequality and moderate deviations for weakly dependent sequences,'' \emph{Probab. Theory Relat. Fields} \textbf{151}(3--4), 435--474 (2011).

\bibitem{Berbee1979}
H.~C.~P.~Berbee, \emph{Random Walks with Stationary Increments and Renewal Theory}, Mathematical Centre Tracts, vol.~112. Amsterdam: Mathematisch Centrum, 1979.

\bibitem{Rio2017}
E.~Rio, \emph{Asymptotic Theory of Weakly Dependent Random Processes}, Probability Theory and Stochastic Modelling, vol.~80. Berlin: Springer, 2017.

\bibitem{Vershynin2018}
R.~Vershynin, \emph{High-Dimensional Probability: An Introduction with Applications in Data Science}, Cambridge Series in Statistical and Probabilistic Mathematics, vol.~47. Cambridge: Cambridge Univ. Press, 2018.

\bibitem{BrownLow1996}
L.~D.~Brown and M.~G.~Low, ``Asymptotic equivalence of nonparametric regression and white noise,'' \emph{Ann. Statist.} \textbf{24}(6), 2384--2398 (1996).

\bibitem{RaoMitra1971}
C.~R.~Rao and S.~K.~Mitra, \emph{Generalized Inverse of Matrices and Its Applications}. New York: Wiley, 1971.

\bibitem{StoicaMarzetta2001}
P.~Stoica and T.~L.~Marzetta, ``Parameter estimation problems with singular information matrices,'' \emph{IEEE Trans. Signal Process.} \textbf{49}(1), 87--90 (2001).

\bibitem{LehmannCasella1998}
E.~L.~Lehmann and G.~Casella, \emph{Theory of Point Estimation}, 2nd ed., Springer Texts in Statistics. New York: Springer, 1998.

\bibitem{Klaassen1985}
C.~A.~J.~Klaassen, ``On an inequality of Chernoff,'' \emph{Ann. Probab.} \textbf{13}(3), 966--974 (1985).

\bibitem{BobkovLedoux1998}
S.~G.~Bobkov and M.~Ledoux, ``On modified logarithmic Sobolev inequalities for Bernoulli and Poisson measures,'' \emph{J. Funct. Anal.} \textbf{156}(2), 347--365 (1998).

\bibitem{BardenetMaillard2015}
R.~Bardenet and O.-A.~Maillard, ``Concentration inequalities for sampling without replacement,'' \emph{Bernoulli} \textbf{21}(3), 1361--1385 (2015).

\bibitem{HansonWright1971}
D.~L.~Hanson and F.~T.~Wright, ``A bound on tail probabilities for quadratic forms in independent random variables,'' \emph{Ann. Math. Statist.} \textbf{42}(3), 1079--1083 (1971).

\bibitem{RudelsonVershynin2013}
M.~Rudelson and R.~Vershynin, ``Hanson--Wright inequality and sub-Gaussian concentration,'' \emph{Electron. Commun. Probab.} \textbf{18}, paper no.~82, 1--9 (2013).

\end{thebibliography}
